---
title: "geocompr: general notes"
output: html_notebook
---

General notes of things I want to remember from https://github.com/bogsnork/learn_geocompr


```{r}
library(sf)
library(raster)

library(spData)        # load geographic data
#install.packages("spDataLarge", repos = "https://nowosad.github.io/drat/", type = "source")
library(spDataLarge)   # load larger geographic data
```


# Chapter 2: Geographic Data in R


## simple features `sf`

* open standard for vector data
* `sf` supersedes `sp`, `rgeos`, `rgdal`

```{r}
vignette(package = "sf")
vignette("sf1")
```

```{r}
names(world)
head(world)
```

```{r}
plot(world, max.plot = 10)
```


```{r}
summary(world["lifeExp"])
```


```{r}
world_mini = world[1:2, 1:3]
world_mini
```

## Converting between `sf` and `sp`

```{r}
library(sp)
world_sp = as(world, Class = "Spatial")
world_sp
```


```{r}
world_sf = st_as_sf(world_sp)
world_sf
```


## 2.2.3 Basic map making

```{r}
plot(world[3:6])
plot(world["pop"])
```


You can overlay using `add = TRUE`

```{r}
world_asia = world[world$continent == "Asia", ]
asia = st_union(world_asia)

plot(world["pop"], reset = FALSE)
plot(asia, add = TRUE, col = "red")
```


```{r}
plot(world["continent"], reset = FALSE)
cex = sqrt(world$pop) / 10000
world_cents = st_centroid(world, of_largest = TRUE)
plot(st_geometry(world_cents), add = TRUE, cex = cex)
```

```{r}
india = world[world$name_long == "India", ]
plot(st_geometry(india), expandBB = c(0, 0.2, 0.1, 1), col = "gray", lwd = 3)
plot(world_asia[0], add = TRUE)
```


## 2.2.6 Simple feature geometries (sfg)

```{r}
point_sfg <- st_point(c(5, 2, 3))              # XYZ point

class(point_sfg)

st_geometry_type(point_sfg)
```

```{r}
## POLYGON with a hole
polygon_border = rbind(c(1, 5), c(2, 2), c(4, 1), c(4, 4), c(1, 5))
polygon_hole = rbind(c(2, 4), c(3, 4), c(3, 3), c(2, 3), c(2, 4))
polygon_with_hole_list = list(polygon_border, polygon_hole)
polygon_with_hole = st_polygon(polygon_with_hole_list)
plot(polygon_with_hole)
```




## 2.2.7 Simple feature columns (sfc)

```{r}
# sfc POINT
point1 = st_point(c(5, 2))
point2 = st_point(c(1, 3))
points_sfc = st_sfc(point1, point2)

points_sfc

class(points_sfc)
st_geometry_type(points_sfc)
```

```{r}
# sfc POLYGON
polygon_list1 = list(rbind(c(1, 5), c(2, 2), c(4, 1), c(4, 4), c(1, 5)))
polygon1 = st_polygon(polygon_list1)

polygon_list2 = list(rbind(c(0, 2), c(1, 2), c(1, 3), c(0, 3), c(0, 2)))
polygon2 = st_polygon(polygon_list2)

polygon_sfc = st_sfc(polygon1, polygon2)
st_geometry_type(polygon_sfc)

plot(polygon_sfc)
```


```{r}
# sfc MULTILINESTRING
multilinestring_list1 = list(rbind(c(1, 5), c(4, 4), c(4, 1), c(2, 2), c(3, 2)), 
                            rbind(c(1, 2), c(2, 4)))
multilinestring1 = st_multilinestring((multilinestring_list1))
multilinestring_list2 = list(rbind(c(2, 9), c(7, 9), c(5, 6), c(4, 7), c(2, 7)), 
                            rbind(c(1, 7), c(3, 8)))
multilinestring2 = st_multilinestring((multilinestring_list2))
multilinestring_sfc = st_sfc(multilinestring1, multilinestring2)
st_geometry_type(multilinestring_sfc)
```
It is also possible to create an sfc object from sfg objects with different geometry types:

```{r}
# sfc GEOMETRY
point_multilinestring_sfc = st_sfc(point1, multilinestring1)
st_geometry_type(point_multilinestring_sfc)
plot(point_multilinestring_sfc)
```

All geometries in an sfc object must have the same CRS. 

```{r}
st_crs(points_sfc)

# EPSG definition
points_sfc_wgs = st_sfc(point1, point2, crs = 4326)
st_crs(points_sfc_wgs)
st_crs(world)
plot(points_sfc_wgs)
```

## 2.2.8 The sf class

```{r}
lnd_point = st_point(c(0.1, 51.5))                 # sfg object
lnd_geom = st_sfc(lnd_point, crs = 4326)           # sfc object
lnd_attrib = data.frame(                           # data.frame object
  name = "London",
  temperature = 25,
  date = as.Date("2017-06-21")
  )
lnd_sf = st_sf(lnd_attrib, geometry = lnd_geom)    # sf object

lnd_sf
```


```{r}
st_coordinates(lnd_sf)
```

```{r}
st_proj_info(type = "datum")

st_proj_info(type = "proj")
```

## 2.4.3 CRSs in R

```{r}
vector_filepath = system.file("vector/zion.gpkg", package = "spDataLarge")
new_vector = st_read(vector_filepath)
st_crs(new_vector) # get CRS
new_vector = st_set_crs(new_vector, 4326) # set CRS

```


```{r}
plot(new_vector)
```



```{r}
raster_filepath = system.file("raster/srtm.tif", package = "spDataLarge")
new_raster = raster(raster_filepath)
projection(new_raster) = "+proj=utm +zone=12 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 
                            +units=m +no_defs" # set CRS

plot(new_raster)
```

# Chapter 3 Attribute data operations

```{r}
library(sf)
library(raster)
library(dplyr)
library(stringr) # for working with strings (pattern matching)
library(tidyr) # for unite() and separate()
library(spData)

```


# Chapter 3 Attribute data operations

```{r}
library(sf)
library(raster)
library(dplyr)
library(stringr) # for working with strings (pattern matching)
library(tidyr) # for unite() and separate()

library(spData)
```

```{r}
methods(class = "sf")
```

?? what does the ST stand for?

```{r}
class(world)
world_df = st_drop_geometry(world)
class(world_df)
```

```{r}
sel_area = world$area_km2 < 10000
summary(sel_area) # a logical vector
#>    Mode   FALSE    TRUE 
#> logical     170       7
small_countries = world[sel_area, ]
```


```{r}
small_countries = world[world$area_km2 < 10000, ]
```

```{r}
world1 = dplyr::select(world, name_long, pop)
names(world1)
```


```{r}
# all columns between name_long and pop (inclusive)
world2 = dplyr::select(world, name_long:pop)
world3 = dplyr::select(world, -subregion, -area_km2)
```

```{r}
world4 = dplyr::select(world, name_long, population = pop)
names(world4)
#> [1] "name_long"  "population" "geom"
```

```{r}
world5 = world[, c("name_long", "pop")] # subset columns by name
names(world5)[names(world5) == "pop"] = "population" # rename column manually
```

```{r}
# create throw-away data frame
d = data.frame(pop = 1:10, area = 1:10)
# return data frame object when selecting a single column
d[, "pop", drop = FALSE] # equivalent to d["pop"]
select(d, pop)
# return a vector when selecting a single column
d[, "pop"]
pull(d, pop)
```

```{r}
# data frame object
world[, "pop"]
# vector objects
world$pop
pull(world, pop)
```

```{r}
slice(world, 3:5)
```

```{r}
# Countries with a life expectancy longer than 82 years
world6 = filter(world, lifeExp > 82)
```

```{r}
world7 = world %>%
  filter(continent == "Asia") %>%
  dplyr::select(name_long, continent) %>%
  slice(1:5)
```

```{r}
world8 = slice(
  dplyr::select(
    filter(world, continent == "Asia"),
    name_long, continent),
  1:5)
```

## Vector attribute aggregation 
```{r}
world_agg1 = aggregate(pop ~ continent, FUN = sum, data = world, na.rm = TRUE)
class(world_agg1)
#> [1] "data.frame"
```

```{r}
world_agg2 = aggregate(world["pop"], by = list(world$continent),
                       FUN = sum, na.rm = TRUE)
class(world_agg2)
plot(world_agg2)
#> [1] "sf"         "data.frame"
```

```{r}
world_agg3 = world %>%
  group_by(continent) %>%
  summarize(pop = sum(pop, na.rm = TRUE))

plot(world_agg3)
```

```{r}
world %>% 
  dplyr::select(pop, continent) %>% 
  group_by(continent) %>% 
  summarize(pop = sum(pop, na.rm = TRUE), n_countries = n()) %>% 
  top_n(n = 3, wt = pop) %>%
  arrange(desc(pop)) %>%
  st_drop_geometry()
```

## 3.2.3 Vector attribute joining

```{r}
world_coffee = left_join(world, coffee_data)
#> Joining, by = "name_long"
class(world_coffee)
#> [1] "sf"         "tbl_df"     "tbl"        "data.frame"
```

```{r}
names(world_coffee)
#>  [1] "iso_a2"                 "name_long"              "continent"             
#>  [4] "region_un"              "subregion"              "type"                  
#>  [7] "area_km2"               "pop"                    "lifeExp"               
#> [10] "gdpPercap"              "geom"                   "coffee_production_2016"
#> [13] "coffee_production_2017"
plot(world_coffee["coffee_production_2017"])
```

```{r}
coffee_renamed = rename(coffee_data, nm = name_long)
world_coffee2 = left_join(world, coffee_renamed, by = c(name_long = "nm"))
```


```{r}
world_coffee_inner = inner_join(world, coffee_data)
#> Joining, by = "name_long"
nrow(world_coffee_inner)
#> [1] 45
```

```{r}
setdiff(coffee_data$name_long, world$name_long)
#> [1] "Congo, Dem. Rep. of" "Others"
```

```{r}
coffee_data$name_long[grepl("Congo,", coffee_data$name_long)] = 
  str_subset(world$name_long, "Dem*.+Congo")
world_coffee_match = inner_join(world, coffee_data)
#> Joining, by = "name_long"
nrow(world_coffee_match)
#> [1] 46
```


```{r}
aggregate
```


# Chapter 8 Making maps with R

```{r}
i
```



