---
title: "chapter 3
exercises"
output: html_notebook
---
## 2.6 Exercises

```{r}
library(sf)
library(spData)
library(spDataLarge)
library(raster)

library(stringr) # for working with strings (pattern matching)
library(tidyr)   # for unite() and separate()
```


### Exercise 1
Create a new object called us_states_name that contains only the NAME column from the us_states object. What is the class of the new object and what makes it geographic?

```{r}
data(us_states)
data(us_states_df)
```


```{r}
us_states

us_states_name <- us_states["NAME"]
class(us_states_name)
```

# 2  Select columns from the us_states object which contain population data. Obtain the same result using a different command (bonus: try to find three ways of obtaining the same result). Hint: try to use helper functions, such as contains or starts_with from dplyr (see ?contains).

```{r}
us_states[c("total_pop_10", "total_pop_15")]
us_states %>% select(total_pop_10, total_pop_15)
us_states %>% select_at(vars(contains("pop")))
us_states %>% select_at(vars(starts_with("total")))
```
# 3
Find all states with the following characteristics (bonus find and plot them):  

* Belong to the Midwest region.
* Belong to the West region, have an area below 250,000 km2 and in 2015 a population greater than 5,000,000 residents (hint: you may need to use the function units::set_units() or as.numeric()).
* Belong to the South region, had an area larger than 150,000 km2 or a total population in 2015 larger than 7,000,000 residents.

```{r}
names(us_states)
```
* Belong to the Midwest region.

```{r}
us_states %>% filter(REGION == "Midwest") %>% plot()


```
* Belong to the West region, have an area below 250,000 km2 and in 2015 a population greater than 5,000,000 residents (hint: you may need to use the function units::set_units() or as.numeric()).
```{r}
str(us_states)
us_states %>%
  filter(REGION == "West" & 
           #as.numeric(AREA) < 250000 &
           AREA < set_units(250000, "km^2") &
           total_pop_15 > 5000000)

```
* Belong to the South region, had an area larger than 150,000 km2 or a total population in 2015 larger than 7,000,000 residents.

```{r}
us_states %>%
  filter(REGION == "South" & as.numeric(AREA) < 150000 & total_pop_15 > 7000000) %>%  plot()
```
# 4

What was the total population in 2015 in the us_states dataset? What was the minimum and maximum total population in 2015?

```{r}
us_states %>%
  summarise(total = sum(total_pop_15), 
            min = min(total_pop_15), 
            max = max(total_pop_15)) #%>% plot()

```
#5
How many states are there in each region?



```{r}
us_states %>%
  group_by(REGION) %>%
  summarise(n = n()) 
```
# 6
What was the minimum and maximum total population in 2015 in each region? What was the total population in 2015 in each region?
```{r}
us_states%>%
  group_by(REGION) %>%
  summarise(min = min(total_pop_15), 
            max = max(total_pop_15),
            total_pop_15 = sum(total_pop_15))


aggregate(total_pop_15 ~ REGION, FUN = sum, data = us_states, na.rm = TRUE)
```

# 7

Add variables from us_states_df to us_states, and create a new object called us_states_stats. 

```{r}
us_states_stats <- left_join(us_states, us_states_df, by = c(NAME = "state"))
```


What function did you use and why? 

Left join


Which variable is the key in both datasets? 

NAME and state

What is the class of the new object?
```{r}
class(us_states_stats)
```

#8

us_states_df has two more rows than us_states. How can you find them? (hint: try to use the dplyr::anti_join() function)

```{r}
anti_join(us_states_df, us_states, by = c(state = "NAME"))

```

#9
What was the population density in 2015 in each state? What was the population density in 2010 in each state?

```{r}
us_states_stats %>% names()

us_states_stats %>% 
  mutate(density_10 = total_pop_10/AREA, 
         density_15 = total_pop_15/AREA) %>% 
  select(NAME, density_10, density_15)
```
# 10
How much has population density changed between 2010 and 2015 in each state? Calculate the change in percentages and map them.

```{r}
us_density <- us_states_stats %>% 
  mutate(density_10 = total_pop_10/AREA, 
         density_15 = total_pop_15/AREA) %>%
  mutate(density_change = density_10 - density_15)

plot(us_density["density_change"])
```

# 11
Change the columns' names in us_states to lowercase. (Hint: helper functions - tolower() and colnames() may help.)

```{r}
us_states_lc <- us_states
(names(us_states_lc) <- tolower(names(us_states_lc)))
```

# 12
Using us_states and us_states_df create a new object called us_states_sel. The new object should have only two variables - median_income_15 and geometry. Change the name of the median_income_15 column to Income.

```{r}
us_states_sel <- us_states %>% 
  left_join(us_states_df, by = c(NAME = "state")) %>% 
  select(Income = median_income_15, geometry)

us_states_sel
```

# 13
Calculate the change in median income between 2010 and 2015 for each state. 

Bonus: What was the minimum, average and maximum median income in 2015 for each region? What is the region with the largest increase of the median income?

```{r}
us_states_stats %>% mutate(income_change = median_income_10 - median_income_15) %>% 
  group_by(REGION) %>% 
  summarise(min = min(median_income_15), mean = mean(median_income_15), max = max(median_income_15), max_increase = max(income_change))
```







